---
title: "CAFOs vs Health"
author: "Theodore Wang"
date: "10/6/2020"
output:
  pdf_document: default
  html_document: default
---

# Background

North Carolina is home to numerous concentrated animal feeding operations (CAFO). Previous studies have already shown that CAFOs can cause health effects in polluting air and water; CAFO waste is known to contain harmful chemicals such as methane, ammonia, and hydrogen sulfide.

How do CAFOs affect local community health? I want to focus on these specific health variables: average number of physically unhealthy days, average number of mentally unhealthy days, food environment index, average daily PM2.5 (air pollution), life expectancy, percent frequent physical distress, percent frequent mental distress, and percent poor or fair health.

Can we use previous recorded data on environmental quality (air, water) surrounding CAFOs in order to determine the overall general health of nearby communities in the future?

What steps can we take to address the health issues regarding CAFOs after identifying areas with the highest concentration of hogs?

# Preparation
```{r check_packages, include = FALSE}
packages <- c("janitor", "tidyverse", "readxl")
for (p in packages) {
  if (!requireNamespace(p)) {
    install.packages(p)
  }
}
```

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)
library(janitor)
```

# Importing Data

## Loading in CAFO and County Health Data
```{r load_other_data}
# Load in CAFO data
cafoData <- read_excel("./List_Of_Permitted_Animal_Facilities2019-11-06.xls", sheet = 1, skip = 2)

# Load in county health data, 4th sheet
countyHealthFour <- read_excel("./2020 County Health Rankings North Carolina Data - v1_0.xlsx", sheet = 4, skip = 1)

# Load in county health data, 5th sheet
countyHealthFive <- read_excel("./2020 County Health Rankings North Carolina Data - v1_0.xlsx", sheet = 5, skip = 1)
```

# Tidying and Transforming Data

## CAFO Data
```{r clean_CAFO}
# Dropping NAs
cafoNoNA <- drop_na(cafoData)

# Selecting only two variables from this data set; no need to focus on others
cafoVariables <- cafoNoNA %>% select(c("County Name", "Allowable Count")) %>% clean_names(case = "snake")

# "Collapsing" data rows together; 'aggregate()' will combine repeated rows together to create one sum of CAFO allowable count per county
cafoVariables <- aggregate(allowable_count~county_name, data = cafoVariables, FUN = sum)

head(cafoVariables)
```

## County Health Data
```{r clean_county}
# Selecting the data variables I specified earlier using 'select()'; also renaming certain data columns using 'rename()' and cleaning them with 'clean_names()'
countyVariablesFour <- countyHealthFour %>% select(c("County", "Average Number of Physically Unhealthy Days", "Average Number of Mentally Unhealthy Days", "Average Daily PM2.5", "Food Environment Index", "% Fair or Poor Health",)) %>% rename("Air Quality Score Rating" = "Average Daily PM2.5") %>% rename("County Name" = "County") %>% rename("Physically Unhealthy Days" = "Average Number of Physically Unhealthy Days") %>% rename("Mentally Unhealthy Days" = "Average Number of Mentally Unhealthy Days") %>% rename("Percent Poor Health" = "% Fair or Poor Health") %>% clean_names(case = "snake")

head(countyVariablesFour)

# Selecting the rest of the health variables from the fifth sheet
countyVariablesFive <- countyHealthFive %>% select(c("County", "Life Expectancy", "% Frequent Physical Distress", "% Frequent Mental Distress")) %>% rename("County Name" = "County") %>% rename("Percent Physical Distress" = "% Frequent Physical Distress") %>% rename("Percent Mental Distress" = "% Frequent Mental Distress") %>% clean_names(case = "snake")

head(countyVariablesFive)
```

## Joining Data
```{r join_data}
# Joining together the county variable datasets
countyVariables <- countyVariablesFour %>% left_join(countyVariablesFive, by = c("county_name" = "county_name"))

# Joining together the county variable and CAFO datasets
finalData <- countyVariables %>% left_join(cafoVariables, by="county_name")

# Clean up the final joined dataset; replace the NA values in the CAFO data portion with 0 and drop rows without county names
finalData <- finalData %>% replace_na(list("allowable_count" = 0)) %>% drop_na("county_name")

head(finalData)
```

# Visualizing Data

## Histogram of CAFOs (statewide)
```{r histogram_total, message = FALSE, warning = FALSE}
finalData %>% ggplot(aes(x = allowable_count)) + geom_histogram(fill="cyan3") + labs(title = "Histogram of Allowable CAFO Counts") + theme_minimal()
```

# Histogram of CAFOs (county basis)
``` {r histogram_county, message = FALSE, warning = FALSE}
finalData %>% ggplot(aes(x = allowable_count)) + geom_histogram(fill="cyan3") + labs(title = "Histogram of Allowable CAFO Counts in all Counties") + theme_minimal() + facet_wrap(~county_name)
```

# Scatter Plot of CAFO Count vs Mentally Unhealthy Days
```{r scatterPlot_mental, message = FALSE, warning = FALSE}
finalData %>% ggplot(aes(x = allowable_count, y = mentally_unhealthy_days)) + geom_point(color = "darkorchid1", alpha = 0.35) + geom_smooth() + labs(title = "Mentally Unhealthy Days vs Allowable CAFO Count of All Counties") + theme_minimal() + theme(axis.text.y = element_text(size = 6))
```

# Scatter Plot of CAFO Count vs Physically Unhealthy Days
```{r scatterPlot_physical, message = FALSE, warning = FALSE}
finalData %>% ggplot(aes(x = allowable_count, y = physically_unhealthy_days)) + geom_point(color = "darkorchid1", alpha = 0.35) + geom_smooth() + labs(title = "Mentally Unhealthy Days vs Allowable CAFO Count of All Counties") + theme_minimal() + theme(axis.text.y = element_text(size = 6))
```
